An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-09-30 nips_2016.tex
Touch ups to Section 6.1, now will read the Poisson process in detail

The deconvolution approach can be used to construct a correction variable
$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
saved offline. As we have shown above, errors in the ``limit'' function
propagate directly through as errors in the acceptance test.  To achieve a test
error of say $1e-6$ (close to single floating point precision), we need a $y'$
spacing of $1e-6$. It should not be necessary to tabulate values all the way to
$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
resolution is 10,000, this leads to a table with about 1 billion values, which
can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
in $y$, it should be possible to achieve similar accuracy with a much smaller
table. We leave further analysis and experiments with $g(x,y)$ as future work.
%{\color{blue} Daniel: I will look at the code to understand this paragraph
%better because right now I am confused.}
% JFC: the code doesnt implement this yet.

@@

We start with a simple Gaussian mixture model, borrowing an experiment
from~\cite{langevin_2011}.  The parameter is 2-D, $\theta =
(\theta_1,\theta_2)$, and the parameter/data generation process is
\begin{equation}\label{eq:data_generation}
    (\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
    0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
\end{equation}
We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
(0,1)$. The original paper sampled 100 data points form this distribution,
and estimated the posterior in the parameters. We are interested in performance
on larger problems and so sampled 1,000,000 points. The posterior
$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with the prior based on the
$\theta$ generation process in Equation~\ref{eq:data_generation}. The larger
number of samples produces a much sharper posterior with two very narrow peaks.
Our goal is to reproduce the original posterior, so we adjust the temperature
to $T=10,000$. Taking logs, we
get the target as shown in the far left of Figure~\ref{fig:gauss_mix_1}.
%% {\color{blue} Daniel: the ${\rm std}(\Delta')$ stuff is outdated; let's figure
%% out a way to change this. We may also have to justify such a high temperature,
%% because this is far from the ``real'' (log) posterior.}

We run MCMC with our MH test using minibatch size $m=100$.  We also
run this using the methods from Korattikara
2014~\cite{cutting_mh_2014} and Bardenet
2014~\cite{icml2014c1_bardenet14}. For~\cite{cutting_mh_2014}, we use
$m=100$ and increment minibatches by 100 as recommended for that test,
and for~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the
minibatch size geometrically with a ratio of $\gamma = 1.5$ as
recommended. These tests use different minibatch size updates because
of their different approaches to repeated
testing. \cite{icml2014c1_bardenet14} uses straightforward
discounting, allocating a fraction of the test error $\delta$ to each
minibatch test (call these subtests). Its important to bound the
number of subtests therefore, so that the tolerable error per subtest is
not too small. \cite{cutting_mh_2014} uses a more aggressive repeated testing
approach, explicitly modeling correlation between tests and allowing many
more tests for a given $\delta$.

The tolerance for making a decision in~\cite{cutting_mh_2014} is
$\epsilon=0.005$, and the error bound control parameters
in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.
all three use the same random walk proposer
with covariance $\Sigma={\rm diag}(0.3, 0.3)$. Since the proposer
generates a random walk, all shaping of distribution of the samples
is due to the M-H test. All methods are run 5000
times to collect 5000 samples.

@@

In this paper, we have derived a new MH test for minibatch MCMC methods. We demonstrated how a
simple deconvolution process allows us to use a minibatch approximation to the full data tests. We
experimentally show the benefits of our test on Gaussian mixtures and logistic regression.
Straightforward directions for future work include running more experiments with a
particular focus on investigation of the variance precondition.  More elaborate extensions include
combining our results with Hamiltonian Monte Carlo methods, providing a recipe for how to use our
algorithm (following the framework of~\cite{sgmcmc_2015}), or integrating parallel
MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.

>>>>

The deconvolution approach can be used to construct a correction variable
$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
saved offline. As we have shown above, errors in the ``limit'' function
propagate directly through as errors in the acceptance test.  To achieve a test
error of $10^{-6}$ (close to single floating point precision), we need a $y'$
spacing of $10^{-6}$. It should not be necessary to tabulate values all the way to
$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
resolution is 10,000, this leads to a table with about 1 billion values, which
can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
in $y$, it should be possible to achieve similar accuracy with a much smaller
table. We leave further analysis and experiments with $g(x,y)$ as future work.
%{\color{blue} Daniel: I will look at the code to understand this paragraph
%better because right now I am confused.}
% JFC: the code doesnt implement this yet.

@@

We start with a simple Gaussian mixture model, borrowing an experiment
from~\cite{langevin_2011}.  The parameter is 2-D, $\theta =
(\theta_1,\theta_2)$, and the parameter/data generation process is
\begin{equation}\label{eq:data_generation}
    (\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
    0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
\end{equation}
We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$.  We fix $\theta =
(0,1)$. The original paper sampled 100 data points form this distribution,
and estimated the posterior in the parameters. We are interested in performance
on larger problems and so sampled 1,000,000 points. The posterior
$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with the prior based on the
$\theta$ generation process in Equation~\ref{eq:data_generation}. The larger
number of samples produces a much sharper posterior with two very narrow peaks.
Our goal is to reproduce the original posterior, so we adjust the temperature
to $T=10,000$. Taking logs, we get the target as shown in the far left of
Figure~\ref{fig:gauss_mix_1}.

We run MCMC with our MH test using minibatch size $m=100$.  We also
run this using the methods from~\cite{cutting_mh_2014} (``Korattikara 2014'')
and~\cite{icml2014c1_bardenet14} (``Bardenet 2014'').
For~\cite{cutting_mh_2014}, we use $m=100$ and increment minibatches by 100 as
recommended for that test, and for~\cite{icml2014c1_bardenet14}, we use $m=100$
and increase the minibatch size geometrically with a ratio of $\gamma = 1.5$ as
recommended. These tests use different minibatch size updates because of their
different approaches to repeated testing. \cite{icml2014c1_bardenet14} uses
straightforward discounting, allocating a fraction of the test error $\delta$ to
each minibatch test (call these subtests). Its important to bound the number of
subtests therefore, so that the tolerable error per subtest is not too small.
\cite{cutting_mh_2014} uses a more aggressive repeated testing approach,
explicitly modeling correlation between tests and allowing many more tests for a
given $\delta$.

The tolerance for making a decision in~\cite{cutting_mh_2014} is
$\epsilon=0.005$, and the error bound control parameters
in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta = 0.01$.  All methods
(including ours) use the same random walk proposer with covariance $\Sigma={\rm
diag}(0.3, 0.3)$.  Since the proposer generates a random walk, all shaping of
distribution of the samples is due to the M-H test. All methods are run 5000
times to collect 5000 samples.

@@

In this paper, we have derived a new MH test for minibatch MCMC methods. We
demonstrated how a simple deconvolution process allows us to use a minibatch
approximation to the full data tests. We experimentally show the benefits of our
test on Gaussian mixtures and logistic regression.  Straightforward directions
for future work include running more experiments with a particular focus on
investigation of the variance precondition.  More elaborate extensions include
testing on deep neural networks and combining our results with Hamiltonian Monte
Carlo methods, providing a recipe for how to use our algorithm (following the
framework of~\cite{sgmcmc_2015}), or integrating parallel
MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.
